### 2016-09-27 rebuttal.tex

\documentclass[answers,12pt]{exam}
\usepackage{xcolor}

\usepackage{xr-hyper}
\usepackage{xr}
\externaldocument[P-]{paper}

\usepackage{hyperref}
\usepackage{upquote}
\usepackage{textcomp}
\usepackage[T1]{fontenc}

@@

\pagebreak

\section{Comments from Editor}

\begin{questions}
\question I found the \url{http://live.sympy.org/} site very convenient for trying things
out in SymPy, but it is only mentioned in the supplement, which might be
overlooked by readers. I suggest moving that section into the main paper, as
it provides a great way to play along with the examples while reading the
paper.
\begin{solution}
\label{editorpoint1}
This sentence was added to the introduction:
\begin{quote}
All examples could be tested on the \href{http://live.sympy.org}{SymPy
Live} instance, that is an online Python shell, which uses the Google
App Engine to execute SymPy code.
\end{quote}
\end{solution}
\question The Basic Usage section omits what I think is an important point:
how does one distinguish between evaluating exp(1) in Python and exp(1) in
SymPy? In other words, how are symbolic constants specified?
\begin{solution}
  Both Python standard library and SymPy have exp callable, so it depends
  which symbols are imported in your global namespace: the \texttt{exp}
  function from the \texttt{math} module or the \texttt{exp} class from SymPy.
  In the later case a special symbolic constant will be produced.

We have added a sentence to the section~\ref{P-sec:core}, that shows symbolic
constants as an example of leaf nodes:
\begin{quote}
Symbols or symbolic constants, like $e$ or $\pi$, are other examples of
leaf nodes.
\begin{verbatim}
>>> exp(1)
E
>>> exp(1).args
()
>>> x.args
()
\end{verbatim}
\end{quote}
\end{solution}
\question The first
thing I tried after seeing Table 2 (simplication functions) did not work:

@@

It is also worth noting that none of the referee materials we received mention MPFR\@.
\end{solution}
\question Section 4.1: please state whether the syntax for
functions in mpmath is identical, or not, to that in Sympy.
\begin{solution}
The mpmath library is not a SymPy submodule, but a separate library,
which we have indicated in the text. Most function names in mpmath are the same as SymPy, but it is not true in every case. We have
added an example to the text showing an instance where they diverge
(summations).
\end{solution}
\question Unless I have
missed something, a weakness of the paper is that speed is mentioned only
twice, on lines 496 and 666. My assumption is that SymPy is slow compared with
its commercial competitors. Please comment on the speed issue, and not just in
the conclusions.
\begin{solution}
A new paragraph titled ``Speed'' was added to the manuscript, which discusses the
speed issue and lays out a path forward. The conclusion was simplified.
\end{solution}
\question Editors are needed for [20].
\begin{solution}
We have done this.
\end{solution}
\end{questions}

\section{Reviewer 1}
\subsection{Basic reporting}
\begin{questions}
\question Generally good. A part that confused me is the assertion (footnote 3) that ``If
A and B are Symbols created with commutative=False then SymPy will keep A·B
and B·A distinct.'' Does that mean that BOTH of them must be created this way,
and that A and x (if x is created normally) will commute? Is there any way to
declare commutators? How does one guarantee that other pieces of code, e.g.
Gaussian elimination, respect non-commutativity?
\begin{solution}
  We have clarified the footnote. Both expressions must be set as
  \texttt{commutative=False}. Internally, in \texttt{Mul}, the ``commutative
  part'' of an expression is pulled out to the front and canonically ordered,
  and the ``noncommutative part'' is not reordered.

Commutators are implemented in the \texttt{sympy.physics.quantum} module and
are discussed in section~\ref{P-sec:quantum}.

@@

\begin{solution}
We have used \verb|import *| on purpose, as we felt that explicitly importing
names would unnecessarily distract from the content of the paper, but the
reviewer is absolutely right that this is bad practice for actual code. We
have added a footnote:
\begin{quote}\texttt{import *} has been used here to aid the
  readability of the paper, but is best to avoid such wildcard import
  statements in production code, as they make it unclear which names are
  present in the namespace. Furthermore, imported names could clash with
  already existing imports from another package. For example, SymPy, the
  standard Python \texttt{math} library, and NumPy all define the \texttt{exp}
  function, but only the SymPy one will work with SymPy symbolic expressions.
\end{quote}

We have also removed the reference to \texttt{from mpmath import *} from line 485.
\end{solution}

\question Lines 151--153
Minor comment: srepr is a useful command. As someone who also uses Mathematica, I wonder if there is a sympy version of the TreeForm command which produces a visualisation of the expression tree, or maybe an output format that I could pass to a graph library for visualisation?

\begin{solution}
We have added a footnote about the \texttt{dotprint} function (footnote~\ref{P-note:dotprint}),
which outputs expressions in the dot format that can be rendered with Graphviz.
\end{solution}

@@

% TODO For projects that use SymPy
% TODO: Make intro sentence longer.
% TODO: Make it into a table
% TODO: Project name, domain, description

The ``Other Projects that
Depend on SymPy'' section (which has been renamed from ``Other Projects that
Use SymPy'') is {\color{red}TODO: Add reasons here, or else move this to the
paper as the reviewer requests.}

The ``Comparison with Mathematica'' section has been placed in the supplement
because we do not feel that a treatment in the main manuscript would be fair
without an equal comparison to other principle computer algebra systems, such
as Maple, SageMath, and Maxima. Unfortunately, our authorship is inexpert in
these systems, so a complete and fair comparison is impossible.

We deem all the sections in section 3 (now section~\ref{P-sec:features}) to be important core
components of the library. Regarding the supplement sections (previously) 5,
7, and 9---``Sets'', ``Category Theory'', and ``SymPy Live'',
respectively:
\begin{itemize}
\item In SymPy 1.0, the sets module is still in
a relatively infant stage. It is mentioned already in conjunction with the
\texttt{solveset} function. We have added some additional text to the solvers
subsection regarding \texttt{sympy.sets}.

@@

\item The ``Statistics'' module is not a core module and is independent (no
  other SymPy submodules depend on it).
\end{itemize}

The ``Architecture'' section is a key component of the paper. The goal of the
paper is to discuss the architecture and core features of SymPy. It therefore
would be inappropriate to move this section to the supplement. We have
followed the reviewer's advice and moved the architecture section to the end
of the paper, to make the paper easier to follow.
\end{solution}
\end{questions}

\begin{questions}

\question Explaining what Python is (lines 70--72) should go before talking about SymPy as a ``CAS written in Python''. Further, the paper assumes a moderate familiarity with Python (for example, Python's console, OOP, and exceptions), and this should be specified. There should be a short note on the used Python console (\texttt{>>>} is the prompt, with the results of computation following immediately in the lines after it). The citation [25] from line 65 should be moved next to ``Python'' in line 70.
\begin{solution}
The citation was moved and a footnote~\ref{P-note:prompt} for the Python interactive prompt was
  added where it was used first time. We have also added a footnote indicating
  a moderate familiarity with Python is assumed.
\end{solution}

@@

\question In line 184, ``symbols are'' should be replaced by ``\texttt{t} is'' (the general rule is already given in line 179).
\begin{solution}
We have done this.
\end{solution}

\question The code in line 210 should be made into its own line (like a displaymath formula), for typesetting reasons and better readability.
\begin{solution}
We now use a verbatim environment here.
\end{solution}

\question The footnote 4 should be moved from line 221 to line 222, right after ``Basic''.
\label{rev3point16}
\begin{solution}
We have made the suggested change.
\end{solution}

\question The part ``which defines some basic methods for symbolic expression trees'' should be removed from line 222, as it was already given in line 130.
\begin{solution}
This part was removed and footnote moved to the mentioned
text in the section~\ref{P-sec:core}, ``The Core''.
\end{solution}

@@

\question The title ``Features'' in line 276 is ambiguous, as a ``feature''
has no precise meaning in Python (or even software libraries in general). It
should be replaced by ``Packages and modules'' or a similar more precise
wording. The same goes for ``feature'' in most other places in the paper (for
example, the caption of Table 1.\ and line 495). It would be very useful to also include actual names of the packages/modules in Table 1., as well as in any section covering those packages/modules.
\begin{solution}
IEEE 829 defines the term feature as ``a distinguishing
characteristic of a software item (e.g., performance, portability,
or functionality).''  We believe, that this definition fits our needs.

Table 1 was extended to include submodule names.
\end{solution}

\question The sets support listing is unnaturally split in two by the ``This includes\ldots'' sentence which would fit better in parentheses.
\begin{solution}
This suggestion was implemented.
\end{solution}

\question Line 355, add a sentence explaining that in SymPy \texttt{str == repr}, because in Python \texttt{repr} is used to get an unambiguous valid Python code representation, while the return value of \texttt{str} is meant to be human-readable.
\begin{solution}
This footnote was added:
\begin{quote}
  Many Python libraries distinguish the \texttt{str} form of an object, which
  is meant to be human-readable, and the \texttt{repr} form, which is mean to
  be valid Python that recreates the object. In SymPy, \texttt{str(expr) ==
    repr(expr)}. In other words, the string representation of an expression is
  designed to be compact, human-readable, and valid Python code that could be
  used to recreate the expression. As it was noted in section~\ref{P-sec:core},
  the \texttt{srepr} function prints the exact, verbose form of an expression.
\end{quote}
\end{solution}

### 2016-09-27 architecture.tex
Add automatic cross-references to the paper (in rebuttal)
See more at:
http://tex.stackexchange.com/questions/3060/referencing-labels-from-other-files

XXX: Section and footnotes numbers are correct, but my pdf viewer
(evince) doesn't open a new document if I click on them.  I hope,
this is a bug.

A useful way to view an expression tree is using the \texttt{srepr} function, which
returns a string representation of an expression as valid Python code\footnote{
The \texttt{dotprint} function from the \texttt{sympy.printing.dot} submodule
prints output to dot format, which can be rendered with Graphviz to
visualize expression trees graphically.}
with all the nested class constructor calls to create the given expression.
\begin{verbatim}
>>> srepr(expr)
"Add(Mul(Symbol('x'), Symbol('y')), Integer(2))"
\end{verbatim}

>>>>

A useful way to view an expression tree is using the \texttt{srepr} function, which
returns a string representation of an expression as valid Python code\footnote{
\label{note:dotprint}
The \texttt{dotprint} function from the \texttt{sympy.printing.dot} submodule
prints output to dot format, which can be rendered with Graphviz to
visualize expression trees graphically.}
with all the nested class constructor calls to create the given expression.
\begin{verbatim}
>>> srepr(expr)
"Add(Mul(Symbol('x'), Symbol('y')), Integer(2))"
\end{verbatim}

### 2016-09-27 introduction.tex
CASs -> CASes
Reviewer 3, comment 3.
Use CAS's (reviewer 3 comment 3)
Reorder section descriptions in the intro
Drop SymPy Live from table of projects, move to intro
* now we have the documentaion reference early
* save space
* we avoid some duplication
Add automatic cross-references to the paper (in rebuttal)
See more at:
http://tex.stackexchange.com/questions/3060/referencing-labels-from-other-files

XXX: Section and footnotes numbers are correct, but my pdf viewer
(evince) doesn't open a new document if I click on them.  I hope,
this is a bug.

Unlike many CASs, SymPy does not invent its own programming language. Python
itself is used both for the internal implementation and end user
interaction. By using the operator overloading functionality of Python, SymPy follows the embedded domain specific language paradigm proposed by Hudak~\cite{dsl-little-languages}.  The exclusive usage of a single programming language makes it easier
for people already familiar with that language to use or develop SymPy.
Simultaneously, it enables developers to focus on mathematics, rather than
language design.  SymPy officially supports Python 2.6, 2.7 and 3.2--3.5.

@@

The remainder of this paper discusses key components of the SymPy library.
Section~\ref{sec:architecture} discusses the architecture of SymPy.
Section~\ref{sec:features} enumerates the features of SymPy and takes a closer
look at some of the important ones. The section~\ref{sec:numerics}
looks at the numerical features of SymPy and its dependency library, mpmath.
Section~\ref{sec:domain_specific} looks at the domain specific physics
submodules for performing symbolic and numerical calculations in classical mechanics
and quantum mechanics. Conclusions and future directions for SymPy are given
in section~\ref{sec:conclusion}.  All examples in this paper use SymPy
version 1.0 and mpmath version 0.19.

The following statement imports all SymPy functions into the global Python
namespace.\footnote{\texttt{import *} has been used here to aid the
  readability of the paper, but is best to avoid such wildcard import
  statements in production code, as they make it unclear which names are
  present in the namespace. Furthermore, imported names could clash with
  already existing imports from another package. For example, SymPy, the
  standard Python \texttt{math} library, and NumPy all define the \texttt{exp}
  function, but only the SymPy one will work with SymPy symbolic expressions.}
From here on, all examples in this paper assume that this statement has been
executed:\footnote{The three greater-than signs denote the user input for the
  Python interactive session, with the result, if there is one, shown on the
  next line.}

@@

All examples could be tested on the \href{http://live.sympy.org}{SymPy
Live} instance, that is an online Python shell, which uses the Google
App Engine to execute SymPy code.

>>>>

Unlike many CAS's, SymPy does not invent its own programming language. Python
itself is used both for the internal implementation and end user
interaction. By using the operator overloading functionality of Python, SymPy follows the embedded domain specific language paradigm proposed by Hudak~\cite{dsl-little-languages}.  The exclusive usage of a single programming language makes it easier
for people already familiar with that language to use or develop SymPy.
Simultaneously, it enables developers to focus on mathematics, rather than
language design.  SymPy officially supports Python 2.6, 2.7 and 3.2--3.5.

@@

The remainder of this paper discusses key components of the SymPy library.
Section~\ref{sec:features} enumerates the features of SymPy and takes a closer
look at some of the important ones. The section~\ref{sec:numerics} looks at
the numerical features of SymPy and its dependency library, mpmath.
Section~\ref{sec:domain_specific} looks at the domain specific physics
submodules for performing symbolic and numerical calculations in classical
mechanics and quantum mechanics. Section~\ref{sec:architecture} discusses the
architecture of SymPy. Conclusions and future directions for SymPy are given
in section~\ref{sec:conclusion}. All examples in this paper use SymPy version
1.0 and mpmath version 0.19.

The following statement imports all SymPy functions into the global Python
namespace.\footnote{\texttt{import *} has been used here to aid the
  readability of the paper, but is best to avoid such wildcard import
  statements in production code, as they make it unclear which names are
  present in the namespace. Furthermore, imported names could clash with
  already existing imports from another package. For example, SymPy, the
  standard Python \texttt{math} library, and NumPy all define the \texttt{exp}
  function, but only the SymPy one will work with SymPy symbolic expressions.}
From here on, all examples in this paper assume that this statement has been
executed:\footnote{\label{note:prompt} The three greater-than signs denote the user input for the
  Python interactive session, with the result, if there is one, shown on the
  next line.}

@@

All examples could be tested on the \href{http://live.sympy.org}{SymPy Live}
instance, that is an online Python shell, which uses the Google App Engine to
execute SymPy code.  The SymPy Live is integrated in the SymPy documentation
system at \href{http://docs.sympy.org}{http://docs.sympy.org}.
